---
title: "Practical Maschine Learning Project"
author: "Christian Kehrle"
date: "27. Dezember 2015"
output: html_document
---

# Backround

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

# Processig and Exploratory Data Analysis

## Data Load

```{r, warning=FALSE}
library(caret, quietly=TRUE)
url_train <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
url_test <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
download.file(url = url_train, destfile = './train.csv',quiet=TRUE,mode='wb')
download.file(url = url_test, destfile = './test.csv',quiet=TRUE,mode='wb')
pml_train <- read.csv(file = 'train.csv',na.strings = c('','NA','#DIV/0!'))
pml_test <- read.csv(file = 'test.csv',na.strings  = c('','NA','#DIV/0!'))
```

## Explorytory Data Analysis

### Training Data 

First Lines of training data set
```{r, eval=FALSE}
head(pml_train)
```
summary on training data set
```{r, eval=FALSE}
summary(pml_train)
```

### Test data

First Lines of test data set
```{r, eval=FALSE}
head(pml_test)
```
summary on test data set
```{r, eval=FALSE}
summary(pml_test)
```

# Modelling and prediction

## Feature enngineering

```{r}
feature_index <- colnames(pml_train)
feature_index <- colnames(pml_train[colSums(is.na(pml_train)) == 0])
feature_index <- feature_index[-c(1:7)]
```

## Cross validation split

To find an optimal model, with the best performance both in Accuracy as well as minimizing Out of Sample Error, the full testing data is split randomly with a set seed with 80% of the data into the training sample and 20% of the data used as cross-validation. When the samples are created, they are sliced by column against the feature set so only the variables of interest are fed into the final model.

For model validation, to measure up accuracy as well as keeping "Out of Sample Error" low, training data itself is split into a training and validation part. This separation is done with a fixed seed and 80% of the data for training and 20% for validation. The splitting is done per partition of the attribute "classe" to ensure availability of every all values in the training set.
```{r}
set.seed(666)
index_train <- createDataPartition(y=pml_train$classe, p=0.80, list=FALSE)
data_train <- pml_train[index_train,feature_index]
data_xval <- pml_train[-index_train,feature_index]
```

## MOdel training and cross validation

```{r,warning=FALSE, results='hide'}
model<- train(classe ~ .,
                data = data_train, 
                method = 'rf', 
                trControl = trainControl(method = "cv", 
                                         number = 4, 
                                         allowParallel = TRUE, 
                                         verboseIter = TRUE))
pred <- predict(model,data_xval)
cm_rf <- confusionMatrix(pred,data_xval$classe)
```

## Variable Importance

```{r}
print(plot(varImp(model, scale = FALSE)))
```

```{r}
cm_rf
```

## Prediction 

```{r}
final_col <- length(colnames(pml_test[]))
colnames(pml_test)[final_col] <- 'classe'
submission <- predict(model,pml_test[,feature_index])
submission
```



